---
title: "R Programming: Advance Analytics in R for Data Science 3"
output: html_notebook
---
You are working on a project for a meteorology bureau. You have been supplied
weather data for 4 cities in the US: Chicago, NewYork, Houston and SanFrancisco.
You are required to deliver the following outputs:
1. A table showing the annual averages of each observed metric for every city
2. A table showing by how much temperature fluctuates each month from min to
max (in %). Take min temperature as the base
3. A table showing the annual maximums of each observed metric for every city
4. A table showing the annual minimums of each observed metric for every city
5. A table showing in which months the annual maximums of each metric were
observed in every city (Advanced)

```{r}
getwd()

setwd("/Users/steph/OneDrive/Documents/R Projeccts")

Chicago<- read.csv("Chicago-F.csv",row.names = 1)
NewYork<- read.csv("NewYork-F.csv",row.names = 1)
Houston<- read.csv("Houston-F.csv",row.names = 1)
SanFrancisco<- read.csv("SanFrancisco-F.csv",row.names = 1)

Chicago
NewYork
Houston
SanFrancisco
is.data.frame(Chicago)
```
Let's convert to matrices
```{r}
Chicago<-as.matrix(Chicago)
NewYork<-as.matrix(NewYork)
Houston<-as.matrix(Houston)
SanFrancisco<-as.matrix(SanFrancisco)
# lets put everything into a list
Weather <- list(Chicago=Chicago,NewYork=NewYork,Houston=Houston,SanFrancisco=SanFrancisco)
Weather[3]
Weather[[3]]
Weather$Houston
summary(Weather$Houston)

```
Example of apply function  in a simple matrix.

The Apply Family is the most powerful set of tools in R.
```{r}
t = matrix(c(27,1,44,63,54,14,42,34,54,111,98,101,12,21,2),ncol=5,nrow=3)
t

noquote("---------------check example below---------------------------")
apply(t,1,mean)             
apply(t,2,max)
apply(t,2,mean)  
```

```{r}
#help(apply)
Chicago
apply(Chicago,1,mean)
#check
mean(Chicago["DaysWithPrecip",])

#analyze one city:

Chicago
apply(Chicago,1,max)
apply(Chicago,1,min)
#for practice
#this does not make much sense but this is good data to compare.
apply(Chicago,2,max)
apply(Chicago,2,min)
#Compare:
apply(Chicago,1, mean)
apply(NewYork,1, mean)
apply(Houston,1, mean)
apply(SanFrancisco,1, mean)
```

```{r}
output<-NULL  #Preparing an empty vector

for(i in 1:5){
  output[i]<-mean(Chicago[i,])
}

output
names(output)<-rownames(Chicago)
output
               #VS

noquote("--------------------------------")
apply(Chicago,1,mean) #only a liine of quote.


#apply can do anything we can do with loops.




```

```{r}
?lapply
Chicago
t(Chicago)
Weather
lapply(Weather,t)
mynewlist<- lapply(Weather,t)
mynewlist



```
Example 2
```{r}
Chicago
rbind(Chicago,NewRow=1:12)
lapply(Weather,rbind,NewRow=1:12)

```
example 3
```{r}
?rowMeans
rowMeans(Chicago) #identical to : apply(Chicago,1,mean)

noquote("-------------------------------")

lapply(Weather,rowMeans)
```
Combining lapply with [] operator
```{r}
#Weather
Weather$Chicago[1,1]
Weather[[1]][1,1]
lapply(Weather,"[",1,1)
#Weather
lapply(Weather, "[",1,)
lapply(Weather, "[", ,3)
```

```{r}
#adding your own functions
lapply(Weather,rowMeans)
noquote("-------------------------------")
lapply(Weather, function(x) x[1,]) #every row from every matrix
noquote("-------------------------------")
lapply(Weather, function(x) x[5,]) #every row from every matrix
noquote("-------------------------------")
lapply(Weather, function(x) x[,1]) #november data
noquote("-------------------------------")
lapply(Weather, function(x) x[,12]) #december data
noquote("-------------------------------")
lapply(Weather, function(o) o[1,]-o[2,]) #december data


```

```{r}
Weather
noquote("-------------------------------")
lapply(Weather,function(z)z[1,]-z[2,])
noquote("-------------------------------")
lapply(Weather,function(z)round((z[1,]-z[2,])/z[2,],2))

```
Average F for July:
```{r}
lapply(Weather,"[",1,7)
noquote("------------------------------")
sapply(Weather,"[",1,7)

```

```{r}
noquote("------------------------------")
lapply(Weather,"[",1,10:12)
noquote("------------------------------")
sapply(Weather,"[",1,10:12)
noquote("------------------------------")
round(sapply(Weather,rowMeans),0)
noquote("------------------------------")
```
Another example
```{r}
lapply(Weather,
       function(z)round((z[1,]-z[2,])/z[2,],2))

round(sapply(Weather,
             function(z)round((z[1,]-z[2,])/z[2,],2)),3)#This function looks better.
```
Nesting lapply function
```{r}
lapply(Weather,rowMeans)
sapply(Weather,rowMeans)
#?rowMeans
Chicago
apply(Chicago, 1, max)


noquote("---------------------------------------")

lapply(Weather,apply, 1, max) # BE  very careful when doing this

sapply(Weather,apply, 1, max) # BE  very careful when doing this
lapply(Weather,function(x)apply(x, 1, max))
sapply(Weather,function(x)apply(x, 1, max))
```
THIS IS THE MOST ADVANCED PART OF THE TUTORIAL
```{r}
#?which.max
which.max(Chicago[1,])
noquote(names
        (which.max
          (Chicago[1,])))

noquote(
  apply(Chicago,1, function(x) names(which.max(x))))


#iterating throughout the whole list

noquote(lapply
        (Weather,
          function(y)apply(Chicago,1, 
                           function(x) names(which.max(x)))))

noquote(
  sapply(Weather,
         function(y)apply(Chicago,1, 
                          function(x) names(which.max(x)))))
```
QUICK RECAP:
1. HOW THE APPLY FAMILY OF FUNCTIONS WORKS.
2.HOW TO RECREATE THE APPLY STATEMENT WITH A FOR() LOOP.
3.WHEN TO USE APPLY(), LAPPLY(), OR SAPPLY.
4. HOW TO COMBINE LAPPLY() WITH[].
5. HOW TO NEST YOUR OWN FUNCTIONS IN APPLY-TYPE FUNCTIONS.
6.HOW TO NEST APPLY-TYPE FUNCTION WITHIN EACH OTHER.
7. HOW TO USE THE WHICH.MAX() AND WHICH.MIN() FUNCTIONS.
```{r}
M = matrix(c(1:50),ncol=5,nrow = 10)
M
M[5,]

```

```{r}

M1 <- rbind(c(100,200,300,400,500), rep(10,5))
M2 <- rbind(1:5, rep(10,5))
MyAwesomeList <- list(M1, M2)
MyAwesomeList






```

