---
title: "R Notebook"
output: html_notebook
---


Code snippets appear like this example, where we simulate 1 million throws
of two six‐sided dice:
```{r}
set.seed(42)
throws<- 1e6
dice<- replicate(2,
                 sample(1:6,throws,replace = TRUE))
table(rowSums(dice))
```

```{r}
a = "Hello world"
print(a)
noquote(a)
paste(a)

```

```{r}
 baskets.of.Granny <- c(12, 4, 4, 6, 9, 3)
sum(baskets.of.Granny)

baskets.of.Geraldine <-
 c(5, 3, 2, 2, 12, 9)
Intro <- "It is amazing! The All Star Grannies scored
a total of"
Outro <- "baskets in the last six games!"
Total.baskets <- baskets.of.Granny +
 baskets.of.Geraldine
Text <- paste(Intro,
 sum(Total.baskets),
 Outro)

print(Text)
noquote(Text)


cat(Text) ##right way
```

x + y y added to x 2 + 3 = 5
x – y y subtracted from x 8 – 2 = 6
x * y x multiplied by y 3 * 2 = 6
x / y x divided by y 10 / 5 = 2
x ^ y x raised to the power y 2 ^ 5 = 32
x %% y remainder of x divided by y (x mod y) 7 %% 3 = 1
x %/% y x divided by y but rounded down
(integer divide)
7 %/% 3 = 2
```{r}
x=2
y=3
x+y
x-y
x*y
x/y
x^y
x%%y
x%/%y

```


```{r}
"+"(x,y)
"-"(x,y)
"*" (x,y)
"/" (x,y)
"^" (x,y)
"%%" (x,y)
"%/%" (x,y)
"min" (x,y)
"max" (x,y)
"mean" (x,y)
"abs"(x)
"sqrt" (x)
"factorial" (x)
"choose" (x,y)
"diff"(x,y)
"prod"(x,y)
```


```{r}

log(x,base = y)
log(c(1:10))
sqrt(c(1:10))
print(123.456,digits = 3)
round(-123.456, digits = -2)
 signif(-123.456, digits = 4)
```


```{r}
print(pi,digits=22)
floor(pi)
ceiling(pi)
trunc(pi)
```


```{r}
funct<- c("is.finite()","is.infinite()","is.nan()","is.na()")

inf<- c("FALSE","TRUE","FALSE","FALSE")
neg.inf<-c("FALSE","TRUE","FALSE","FALSE")
NAN<-c("FALSE","FALSE","TRUE","TRUE")
na<-c("FALSE","FALSE","TRUE","TRUE")
data.frame(funct,inf,neg.inf,NAN,na)

```


```{r}
p<- c(1:10)
p[2]
p[is.na(p)]
p[!is.na(p)]
p[!is.na(2)]
```

x == y Returns TRUE if x exactly equals y
x != y Returns TRUE if x differs from y
x > y Returns TRUE if x is larger than y
x >= y Returns TRUE if x is larger than or exactly equal to y
x < y Returns TRUE if x is smaller than y
x <= y Returns TRUE if x is smaller than or exactly equal to y
x & y Returns the result of x and y
x | y Returns the result of x or y
! x Returns not x
xor( x, y ) Returns the result of x xor y (x or y but not x and y)
```{r}
s<- letters
S<- LETTERS
s[1:7]
S
length(s)
length(S)
s[10]
s[15]
s[5]
s[12]
noquote(s)
noquote(S)
head(s,3)
tail(S,3)
```


```{r}
str(islands)
noquote(names(islands)[1:40])
noquote(names(islands)[12])
b <- noquote(names(islands)[18])
```


```{r}
month.days <- c(31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
names(month.days)<- month.name
month.days
```

Splitting texts
```{r}
pangram <- "The quick brown fox jumps over the lazy dog."
strsplit(pangram, " ")




```
Splitting text
```{r}
txt<-"Do you have to use R on the job, but you don't know much about programming? Relax! Here's a quick and painless guide to using R for all sorts of statistical and data analysis. Through real-world examples and step-by-step exercises you'll learn how to reshape and manipulate data, perform calculations, create graphical representations, and more."
strsplit(txt, " ")
```
Back to normal
```{r}
paste(txt,collapse = " ")
```
```{r}
sort(letters,decreasing = TRUE)

state.name
substr(state.name,start = 1,stop = 6)
```
```{r}
gsub("cheap", "sheep's", "A wolf in cheap clothing.")

x <- c("file_a.csv", "file_b.csv", "file_c.csv")
y <- gsub("file_", "", x)
y
gsub("\\.csv", "", y)
```


```{r}
directions <- c("North", "East", "South", "South")
factor(directions,
       levels = c("North", "East", "South", "West"),
       labels = c("N", "E", "S", "W"))
```

%H Hours as a decimal number (00–23)
%I Hours as a decimal number (01–12)
%M Minutes as a decimal number (00–59)
%S Seconds as a decimal number (00–61)
%p AM/PM indicator
```{r}
apollo <- "July 20, 1969, 20:17:39"
apollo.fmt <- "%B %d, %Y, %H:%M:%S"
xct <- as.POSIXct(apollo, format = apollo.fmt, tz = "UTC")
xct
noquote("__________informal way__________")

format(xct, "%M minutes past %I %p, on %d %B %Y")


#There is an important difference between Date objects and POSIXct or
#POSIXlt objects. If you use a Date object, you add and subtract days; with
#POSIXct and POSIXlt, the operations add or subtract only seconds.
```
```{r}
matrix(c(1:12),ncol = 4,nrow = 3)

noquote("////////////////////////////////")
matrix(c(1:12),ncol = 4,byrow = TRUE)

```

```{r}
a<- c(1:9)
b<-c(10:18)

rbind(a,b)

```

```{r}
baskets.of.Granny <- c(12, 4, 5, 6, 9, 3)
baskets.of.Geraldine <- c(5, 4, 2, 4, 12, 9)
baskets.team <- rbind(baskets.of.Granny, baskets.of.Geraldine)
baskets.team
```
Inverting a matrix
Contrary to your intuition, inverting a matrix is not done by raising it to the
power of –1. As explained in Chapter 6, R normally applies the arithmetic
operators element‐wise on the matrix. So, the command first.matrix^(‐1)
doesn’t give you the inverse of the matrix; instead, it gives you the inverse of
the elements. To invert a matrix, you use the solve() function, like this:
```{r}
square.matrix <- matrix(c(1, 0, 3, 2, 2, 4, 3, 2, 1), ncol = 3)
solve(square.matrix)

my.vector <- 1:24
dim(my.vector)= c(3, 4, 2)

```

```{r}
baskets.df <- as.data.frame(t(baskets.team))

boxplot(baskets.of.Geraldine,baskets.of.Granny)
```
```{r}
employee <- c("John Doe", "Peter Gynn", "Jolie Hope")
salary <- c(21000, 23400, 26800)
startdate <- as.Date(c("2010-11-1", "2008-3-25", "2007-3-14"))
employ.data<- data.frame(employee,salary,startdate)

employ.data

str(employ.data)
```


```{r}
employ.data<- data.frame(employee,salary,startdate,stringsAsFactors = FALSE)

str(employ.data)

```

```{r}

```

```{r}
x <- c(0.458, 1.6653, 0.83112)
percent <- round(x * 100, digits = 1)
result <- paste(percent, "%", sep = "")
print(result)
```

```{r}
x <- 1:5
test <- function(x){
 cat("This is x:", x, "\n")
 rm(x)
 cat("This is x after removing it:", x, "\n")
}

```

```{r}
colnames<-c("col 1", "col 2", "col 3")
rownames<-c("row 1", "row 2", "row 3","row 4")
f<- matrix(c(1:12),ncol = 3)
colnames(f)<-colnames
row.names(f)<-rownames

f  # as is
```


```{r}
as.data.frame(f) #converting it into a dataframe

```

```{r}
names <- c("Valentine Swindle"
"Rex Simser,"
Deloise Langenfeld,
Guillermina Gard,
Kasey Schreffler,
Regan Vivas,
Alisa Melson,
Cheree Munch,
Mercedez Rance
Tynisha Buesing
Landon Puig
Rochel Fleeman
Tabitha Marchi
Luise Posada
Lovie Droz
Mindi Gros
Lenny Corchado
Xuan Bearse
Lupe Mousseau
Trenton Linder
Phil Star
Erminia Irving
Lenard Glessner
Normand Corning
Antone Ketter
Moshe Berryman
Teisha Slowik
Lyndia Kruger
Althea Hardt
Annis Smail
Tonia Krouse
Particia Pender
Alona Been
Wilma Solberg
Bart Palladino
Tempie Derry
Tricia Stanwood
Elmira Bridge
Keesha Pettaway
Laquita Imes
Joni Hilaire
Eliz Jesse
Pura Marcinko
Abraham Pettie
Julieann Rodger
Blake Schulze
Thomasine Kettner
Luetta Gudino
Tania Vandam
Wade Whitham)

as.character(names)
names

```


```{r}
g<-matrix(c(1:25),ncol=5)
g
print("-***********************************")
g<-matrix(c(1:25),ncol=5,byrow = TRUE)
g
```

```{r}
library(calculus)
# numeric matrix
x <- matrix(1:4, nrow = 2)
x
inverse(x)

```

```{r}
## a slowly-convergent integral
integrand <- function(x) {1/((x+1)*sqrt(x))}
integrate(integrand, lower = 0, upper = Inf)
```

```{r}
x<-NULL
x<- 1:10

f<- function(x) {2*x^9-6*x
}
integrate(f,lower=0,upper =90000000000000 )
```

```{r}

f<- function(x) {2*x^9-6*x}
f1<-function(x) {2*x^9-6*x}
f2<-function(x) {2*x^9-6*x}
f3<-function(x) {2*x^9-6*x}
h = matrix(c(f,f1,f2,f3),ncol = 2,nrow = 2)
h
D(f,"x")

```

```{r}
library(ggplot2)
ggplot(data= cars,aes(speed,dist,color=speed))+geom_point()


```


```{r}
ggplot(ToothGrowth,aes(x=dose,y=len, fill=dose))+geom_boxplot()
```

```{r}
p<-ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) +
  geom_boxplot()
p
```

```{r}
ggplot(ToothGrowth, aes(x=dose, y=len)) +
  geom_boxplot(fill='#A4A4A4', color="black")+
  theme_classic()
# Change box plot colors by groups
p<-ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) +
  geom_boxplot()
p
```

```{r}
# Convert the variable dose from a numeric to a factor variable
ToothGrowth
boxplot(ToothGrowth$len~ToothGrowth$dose,
        color="red",xlab = "dose",ylab = "length",main="Tooth growth",grid=TRUE)


```

```{r}
plot(cars, main="cars")
abline(v=c(15,20),col=c("blue","purple"),lty=c(1,2),lwd=c(1,3))
```

```{r}
cars$time<-cars$dist*cars$speed
cars$number=as.double(c(1:50))
cars




```


```{r}
cars

stem(cars$cars.number)

```

```{r}


library(ggplot2)
library(tidyverse)
library(shiny)

stem(faithful$eruptions)
q<- ggplot(faithful,aes(eruptions,waiting,color=eruptions))
q+geom_point()+ggtitle("              Eruptions vs waiting     ")+ theme(plot.background = element_rect(fill = 'green', colour = 'purple'))






```

```{r}
# library
library(ggplot2)
 
# create a data frame
variety=rep(LETTERS[1:7], each=40)
treatment=rep(c("high","low"),each=20)
note=seq(1:280)+sample(1:150, 280, replace=T)
data=data.frame(variety, treatment ,  note)

ggplot(data,aes(variety,note,color=treatment))+geom_boxplot()

ggplot(data,aes(variety,note,fill=treatment))+geom_boxplot()
```

```{r}
ggplot(data,aes(variety,note,color=treatment))+geom_boxplot()+facet_wrap(~treatment)
ggplot(data,aes(variety,note,color=treatment))+geom_boxplot()+facet_wrap(~variety, scale="free")
```


```{r}
library(ggplot2)
ggplot(data,aes(variety,note,color=treatment))+geom_boxplot()+ggtitle("asd")
data
ggplot(data,aes(note,color=treatment))+geom_density()
```


```{r}
library(shiny)

ui <- fluidPage(
  titlePanel(title = "Shiny Exercise"),
  sidebarPanel(width = 4,
    sliderInput("RandNum", label = "choose a number", 
      value = 200, min = 10, max = 100)
  ),
  mainPanel(wideth = 8, plotOutput("scatter"))
)

server <- function(input, output)
{output$scatter <- renderPlot({
  x <- rnorm(input$RandNum)
  y <- rnorm(input$RandNum)
  plot(x, y, xlab = "x", ylab = "y", main = "Scatter plot")
})
 
}

shinyApp(ui = ui, server = server)
```

Usage
integral2(fun, xmin, xmax, ymin, ymax, sector = FALSE, reltol = 1e-6, abstol = 0, maxlist = 5000, singular = FALSE, vectorized = TRUE, ...)

integral3(fun, xmin, xmax, ymin, ymax, zmin, zmax, reltol = 1e-6, ...)
Arguments
fun
function
xmin, xmax
lower and upper limits of x.
ymin, ymax
lower and upper limits of y.
zmin, zmax
lower and upper limits of z.
sector
logical.
reltol
relative tolerance.
abstol
absolute tolerance.
maxlist
maximum length of the list of rectangles.
singular
logical; are there singularities at vertices.
```{r}
library(SimplicialCubature)
S = cbind(c(15,15),c(50,15),c(15,50))
adaptIntegrateSimplex(function(v) v[1]+0.805, S)


```

```{r}
data.matrix(summary(cars))


```

```{r}
library(purrr)
y <- list(0:10, 5.5)
y %>% every(is.numeric)
y %>% every(is.integer)
y %>% some(is.integer)
y %>% none(is.character)
```

```{r}
library(ggplot2)
library(sqldf)
iris<- sqldf("SELECT * FROM iris")

o<- ggplot(iris,aes(Species,Sepal.Length),main="asd")+geom_boxplot(fill = "white", colour = "black",outlier.colour = "red", outlier.shape = 1)
iris
o + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2)

o+ geom_boxplot(aes(colour = Sepal.Length))+ggtitle("Petal and sepal length of three species of iris")+labs(x="Species",y="Sepal length")
```


```{r}
library(ggplot2)
library(plotly)
 #plot
iris_plot<- ggplot(iris,aes(Sepal.Length,Petal.Length,color=Species))+ geom_point()+ggtitle("Petal and sepal length of three species of iris")+labs(x="Sepal Length",y="Petal length") 
#adding a background
iris_plot<- iris_plot + theme(plot.background = element_rect(fill = 'green', colour = 'purple'))
#using plotly
ggplotly(iris_plot)


```

```{r}
p<- ggplot(mpg,aes(class,hwy))
p+geom_boxplot()+theme(plot.background = element_rect(fill = 'yellow', colour = 'blue'))
```

```{r}
p <- ggplot(mpg, aes(class,hwy,fill=class))
p + geom_boxplot()+ggtitle("asd")+labs(x="asd",y="ads")


p <- ggplot(mpg, aes(class,hwy,color=class))
p + geom_boxplot()+ggtitle("asd")+labs(x="asd",y="ads")
```

```{r}
library(ggplot2)

ggplot(mpg,aes(class,displ,color=class))+geom_boxplot()+ggtitle("ads")+labs(x="class",y="display")

```

```{r}
library(ggplot2)
ggplot(mpg,aes(hwy,color=fl))+geom_density()
ggplot(mpg,aes(hwy,color=manufacturer))+geom_density()
ggplot(mpg,aes(manufacturer,color=manufacturer))+geom_density()
```
```{r}
print(x <- "Hello world!")
print(digits=4 , x = 11/7)
print(11/7,digits = 3)

savehistory(file="Joel.Rhistory") #This saaves the entire history.

loadhistory(file="Joel.Rhistory")
```

```{r}
paste("Hello world")
```

```{r}
baskets.of.Granny <- c(12, 4, 4, 6, 9, 3)
baskets.of.Geraldine <-
 c(5, 3, 2, 2, 12, 9)
Intro <- "It is amazing! The All Star Grannies scored
a total of"
Outro <- "baskets in the last six games!"
Total.baskets <- baskets.of.Granny + baskets.of.Geraldine

text <- paste(Intro,sum(Total.baskets),outro)

print((text)) ## This is the wrong way in this case
noquote("*****************************************")
cat(text) #This is the right way.
```
Using mathematical functions
In R, of course, you want to use more than just basic operators. R comes with
a whole set of mathematical functions you’d find on a technical calculator
as well, and then some more. Table 4-2 lists the ones that we think you’ll use
most often, but feel free to go on a voyage of discovery for others. All these
functions are vectorized, so you can use them on complete vectors.
The possibilities of R go far beyond this small list of functions, however. We
cover some of the special cases in the following sections.
```{r}
x=6
y=3
x%%y
x%/%y
log(x,base=y)
abs(x/y)
exp(x)
sqrt(y)
factorial(y)

choose(x,y)
log(1:10)

floor(pi)
ceiling(pi)
trunc(pi)
```
```{r}
"+"(2,3)
"-"(2,3)
"/"(2,3)
"*"(2,3)

is.finite(10^(305:315))
```

```{r}
length(baskets.of.Geraldine)
```
Replacing data in a matrix
```{r}
t <- c(1:5)
t[c(2,5)]<-10
t

t>2
which(t>3) #The answer indicate the positions the data is found.

```
x == y Returns TRUE if x exactly equals y
x != y Returns TRUE if x differs from y
x > y Returns TRUE if x is larger than y
x >= y Returns TRUE if x is larger than or exactly equal to y
x < y Returns TRUE if x is smaller than y
x <= y Returns TRUE if x is smaller than or exactly equal to y
x & y Returns the result of x and y
x | y Returns the result of x or y
! x Returns not x
xor( x, y ) Returns the result of x xor y (x or y but not x and y)
```{r}
The.best <- baskets.of.Geraldine<baskets.of.Granny

which(The.best)
```

Table 4-5 Vector Operations
Function What It Does
sum(x) Calculates the sum of all values in x
prod(x) Calculates the product of all values in x
min(x) Gives the minimum of all values in x
max(x) Gives the maximum of all values in x
cumsum(x) Gives the cumulative sum of all values in x
cumprod(x) Gives the cumulative product of all values in x
cummin(x) Gives the minimum for all values in x from the start
of the vector until the position of that value
cummax(x) Gives the maximum for all values in x from the start
of the vector until the position of that value
diff(x) Gives for every value the difference between that
value and the next value in the vector



Leaflet
```{r}
library(leaflet)
m <- leaflet()%>%
  addTiles() %>%
  addMarkers(lng =-82.577140,lat = 28.082520 )
m%>% addProviderTiles(providers$Stamen.Toner)

```


```{r}
m %>% addProviderTiles(providers$Esri.NatGeoWorldMap)
```

```{r}
library(leaflet)
m<- leaflet()%>%
  addTiles() %>%
  addMarkers(lat =28.082520 ,lng =-82.577140 )%>%
  
  addMiniMap()

  
m
```

```{r}
velocity<- function(dist,time)# variable     
  {
  return(dist/time)
}

```

```{r}

velocity()
```

```{r}
f<- function(x){
  return(5*x^3+1)
}

```

```{r}
library(ggplot2)
x<-c(0,20)
axis<-data.frame(x)
ggplot()+
  stat_function(data=axis,aes(x),fun=f)+
  ggtitle("This is an example of a line graphed")
```


```{r}
g <- NULL
g<- function(x,y)
  return(x^6+x^3-2*x^5*9)

```

```{r}
g(5,6)
x <- c(5,50)
axis<- NULL
axis<-data.frame(x)
ggplot()+
  stat_function(data=axis,aes(x),fun=g)+
  ggtitle("This is an example of a line graphed")
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


