---
title: "Postgres SQL connection and automation"
output: html_notebook
---

db: name will be the database you are wanting to use.

host: will be the host your database is on. Either your localhost or the url to your database.

port: By default the port will be 5432 for postgresql.

user: will be the username for the database you are connecting to
password will be the database password you use when connecting to postgresql Next we can list the tables that are available in our database

CONNECTION SETTINGS
There are six settings needed to make a connection:

Driver - See the Drivers section for setup information.

Server - A network path to the database server.

Database - The name of the schema.

UID - The user’s network ID or server local account.

PWD - The account’s password.

Port - Should be set to 5432


Connecting to the postgres sql database

```{r}

library(odbc)
library(DBI)
library(RPostgreSQL)

password = "Kingjames1611"
DBI::dbDriver('PostgreSQL')
drv=dbDriver("PostgreSQL")
con=dbConnect(drv,dbname="dvdrental",
              host="localhost",
              port=5432,
              user="postgres",
              password=password)

dbListTables(conn = con)
```
using the database
```{r}
library(odbc)
library(DBI)
library(RPostgreSQL)

## Actual code for automating the database

query = '(select rental_id,payment_id
from payment,store
)'


#processing the query

result<- RPostgreSQL::dbGetQuery(conn = con,statement = query)
result
```


Keep in mind that you can also plot the information contained in the database right here in R.

```{r}
plot(result)
summary(result)
```
```{r}
library(odbc)
library(DBI)
library(RPostgreSQL)

dbDisconnect(con)

```

